\documentclass[10pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[backend=bibtex, style=authoryear]{biblatex}

% \addbibresource{~/Dropbox/TAGreview/PhDreferences.bib}

\textwidth=470pt
\oddsidemargin=0pt
\topmargin=0pt
\headheight=0pt
\textheight=650pt
\headsep=0pt

\title{Statement of Research}


\author{Nicholas Santantonio}
\date{\today}


\newcommand{\gxe}{G$\times$E}
\newcommand{\gxg}{G$\times$G}


\begin{document}

\section*{\centering Statement of Research}
\begin{center} Nicholas Santantonio \end{center}

\subsection*{Introduction}

% gxg cows
% cover crop to sequester carbon, 
% htp, 
% weather data
% crop modeling,
% abiotic biotic adaptation
% climate change
% biofuel
% sustainable cropping systems. 


% digital ag
% connect with engineering dept. 
% Fed capacity Hatch funds used only for breeding  

% Lead by Professors Royse Murphy and Don Viands, the Cornell forage breeding program has a rich legacy of providing forage crop varieties to farmers in the Northeast US, as well as collaboration nation-wide, a legacy that I hope to contribute to. 

% Forage research has been drastically under-funded compared to food crops, but renewed interest in sustainable farming has brought them back into the limelight. 

Interest in sustainable agriculture has brought renewed attention to forage crops, which can increase soil nitrogen, sequester carbon, reduce weed populations and generally improve soil health while providing a valuable hay crop. Farms must produce more on less land and with less agronomic inputs, leaving genetics as the primary target for improvement. This need is emphasized in sustainable agriculture where fertilizer and pesticide inputs must be used sparingly and carefully to reduce environmental impact. Marginal lands are a marked exception and a good target for adaptation to reduce competition between forage or biofuel crops and food crops. Implementation of new breeding technologies to increase genetic gain and adaptation has lagged in forages due to the complexity of the genetics, phenotyping and costs associated with evaluation of quality.



\subsection*{Grant 1}

Alfalfa, known as the ``Queen of forages'' for its high protein content, is the most important forage crop in US, outproduced only by maize, soy and wheat. Introgression of beneficial alleles in alfalfa is exceedingly difficult due to high inbreeding depression and the autotetraploid nature of the crop. As an obligate out-crosser, alfalfa must always be bred on a population level, where varieties are released as synthetics to avoid inbreeding and take advantage of population level heterosis. This has limited implementation of marker-based selection because large numbers of individuals must be genotyped and inter-mated to avoid inbreeding in future generations. Further complicating the use of markers is which individual(s) should be selected for genotyping to represent a given variety. Single individuals are not representative of the variety as a whole, and genotyping large numbers of individuals from each variety is costly and restrictive.

To implement genomic selection (GS) in alfalfa for adaptation to sustainable agronomic practice and marginal lands, I propose a new population-level genomic selection framework (popGS). Tissue from many individuals within a variety or breeding line will be bulked for DNA extraction and genotyped together. This will reduce genotyping costs while still allowing the standard genomic prediction framework to operate on allele counts in the population as opposed to allele counts within a single individual. A sequence based marker platform will need to be built to address the need to estimate allele frequencies of varieties, preferably with the ability to track greater than 2 alleles. PopGS will allow for prediction of additive effects for genetic gain, as well as dominance effects to exploit population level heterosis. 

A training set will be developed by genotyping varieties and breeding lines from the Cornell forage breeding population and other sources that have been recently phenotyped. Moving forward, all new entries into the program will be genotyped in this fashion. New populations will be developed with ``blending crosses'' where individuals from phenotyped lines will be planted in relevant proportions and inter-mated to produce the desired population allele frequencies. This allows popGS to work in a fluid fashion, where new varieties are constructed by mixing the best alleles together in optimal frequencies to leverage high additive and non-additive effects in the resulting populations. This methodology may also be used for integration of CRISPR gene edits from single individuals into elite germplasm at high frequencies.

% CRISPR methodology is still in its infancy, but once individual plants can be transformed at all four alleles, crossing into elite germplasm will still be a challenge. Half-sib families will be at best duplex for the CRISPR allele, with further reduction in frequencies due to the inability to intracross or backcross. PopGS can be used to enrich CRISPR allele frequencies until phenotypic evaluation is feasible.

% A modified rhAmpSeq approach will be built to capture sequences likely to differ at more than one site such that multiple alleles can be tracked.  One approach may include building primer sets that are anchored in exons, but span introns, in order to maximize the likelihood of amplification and sequence differences. Specific motifs? THIS PARAGRAPH NEEDS WORK!

The first grant proposal will evaluate the efficacy of popGS, both theoretically and empirically, to be submitted within the first year. Theory and simulation will be supported by genotyping a study that has already been phenotyped, and used as a proof of concept (POC). The POC paper will submitted within the a year as the first milestone of the grant. The diallel consisting of crosses between the nine source populations of alfalfa evaluated at NMSU (Segovia Lerma et al. 2004) is the current target for the POC study, but other trials may also be considered.% We will also genotype varieties that have been previously phenotyped in order to provide a proof of concept (POC) study as the first milestone of the grant. 


\subsection*{Grant 2}

Unlike most other agronomic crops, forages are typically harvested multiple times per year, with three to ten cuts a season depending on latitude. Many forages such as alfalfa are also perennial crops, and require 3 to 4 years of phenotyping to evaluate stand persistence. On top of this, forage quality must be measured to gauge animal nutrition and digestibility, but is expensive to phenotype in both cost and labor. This makes for a heavy phenotypic burden on the breeding program that stands to benefit from the aid of digital ag. 

Forage cover crops are highly effective for soil rejuvenation, but the genetic variability of these processes is largely unknown because looking underground is notoriously difficult. Root phenotypes will be key to modeling carbon and nitrogen cycling, drought response, and nodulation. Tap- and fine-root architecture and turnover phenotypes will be collected through time by sampling 1 meter cores with plots (Santantonio et al. 1977), but strategies to economically select for these traits are desperately needed. %Tap root morphology will be measured using cores within the drill strip row, while fine root structure will be evaluated from cores sampled between drill strips within each plot. Root phenotypes will be taken twice per year, after the first and last harvests of the season, to determine root growth cycles. 

To mitigate the phenotypic burden of multiple years, cuts and the measuring of root traits, hyper spectral imaging data will be collected from drones throughout each growth cycle to produce high throughput phenotypes (HTP). Genetic correlations of low throughput phenotypes (LTPs) and HTPs will be estimated using a training population and used to build genomic prediction models to predict unobserved LTPs. HTPs will also be used to develop population specific growth curves to model development above and below ground under differing biotic and abiotic stressors. Instead of waiting for three year trials to come to completion, fast cycle selection and blending crosses will be performed on a yearly basis using popGS and all available HTPs and LTPs from historical and in-field material. Training models will be updated annually with HTP and LTP information from plots in the field. 

The second grant proposal will evaluate incorporation of HTP into a fast cycle forage breeding program, and will be submitted during the second year. The efficacy of HTP selection will be assessed using several selection schemes, where crosses will be made based on genotype predictions alone, genotype plus HTP, and genotype plus HTP and LTP, along with selection on LTPs alone as a control.   
% (i.e. popGS) of cycle 0 at year 1 (C0Y1), genotype predictions including HTP at year 1, and phenotypic information from year 1 alone. After 1 round of selection, the resulting populations will be planted in the field for evaluation (C1.1). Additional phenotypes from cycle 0, year 2 (C0Y2) and cycle 1, year 1 (C1Y1) will be used to update prediction models for an additional round of selection from both C0 and C1 to produce C1.2 and C2.1 populations, respectively. Three year evaluation of these three generations will be contrasted to determine the merit of HTP assisted popGS. 

\subsection*{Grants 3 and 4}

Forage crops are only one part of an agronomic biological system which includes soil microorganisms, the animals that feed on the forages and the microbiome of those animals. Other than host-pathogen interactions, little attention has been paid to genomic interactions between these organisms (i.e. \gxg), despite an overall notion that they are important. These interactions can be thought of as a special case of the \gxe\ problem, where the covariance of the ``environment'' (e.g. soil microorganisms) can be determined by genotyping that ``environment''. Interactions can then be modeled using standard \gxe\ machinery, where the genetic covariance of the biological combinations is a Kronecker product of their genetic relationships. Two interacting systems will be targeted for research into \gxg: the interaction between legume and soil microbes, and the interaction between animal and feed. 

As a legume, alfalfa and other leguminous forages can form symbiotic relationships with nitrogen fixing bacteria, \emph{Rhizobium}. Unfortunately, the signaling and infection process for nodulation is typically reduced or absent under all but the lowest soil nitrogen levels. Collection, mutagenesis, and transformation of \emph{Rhizobium} may enable the genetic variation necessary for simultaneous \gxg\ selection of host variety and symbiont. Genetic increases in nodulation could allow for the use of less chemical fertilizer, reducing the environmental impact of nitrogen runoff. The use of \gxg\ prediction models will allow for targeted selection of host symbiont combinations for testing that are likely to be beneficial. Whole genome expression of both host and symbiont as an intermediate interaction phenotype may also aid in understanding the genetic variation in signaling between the two species through time, under contrasting levels of available nitrogen. This may help us to understand how to select for changes in gene expression to increase the rate of colonization, as well as nitrogenase and leghemaglobin expression, even when some nitrogen is available in the soil. 

% On the other side of the plant, 

There has been considerable research into nutrition of different feeds and how they interact with the microbiome of the rumen in dairy cows. While measures of forage quality are considered, little is known about the effect of different varieties on the animal and its microbiome. I seek to establish a collaboration with the animal nutrition department, local dairy farmers, and animal genetics companies that serve the Northeast to investigate the potential for synergistic forage and animal breeding. Instead of breeding animals independently of their feed, we can start to breed specialized animals to specialized feeds. While it will take time to establish relationships with animal breeders, dairy and hay farmers, bringing these communities together would set precedent for future long term integrated breeding operations.

% \section{other ideas}

% There is also interest in short term alfalfa crops, which can provide a large percentage of the field health gain of a full alfalfa cycle in only one or two years. This well have to be coupled with selection for high seed production under low planting density, a known a bottle neck in an alfalfa agronomic system. %There is also interest in short term alfalfa crops, which can provide a large percentage of the field health gain of a full alfalfa cycle in only one or two years. 

% \subsection*{Personnel}

% I would like to hire one post doctoral associate and one graduate student for the first year, with the potential for additional graduate students and post docs for the following year depending on grant funding. The post doc will work with me to write the popGS theory and simulation paper. The graduate student will be responsible for development of the population level genotyping platform using seed from the previously phenotyped population we select for the POC, with guidance from myself and the post doc. This will lead to two companion manuscripts, with the first on the genotyping platform, and the second on the results from the popGS POC study. 

% The forage breeding program at Cornell consists of a strong team, whose knowledge and experience will be invaluable to implementing a successful GS breeding program. Additional personnel will be brought on as necessary, including a technician to operate and manage all HTP equipment and activities.

% \subsection*{A forage community}

% One of the challenges is the lack of large phenotypic records for alfalfa in any alfalfa breeding program.

% To encourage collaboration, data sharing and a global initiative, I would like to work toward an initiative to create a North American forage database using an available platform such as T3, or BassavaBase. All tools built by our lab will be made publicly available through the (BRAPI/ galaxy etc?) to allow forage breeding programs throughout the world to make more informed mating decisions. Other programs that produce useful tools for forage breeding will also be encouraged to link them to the $\alpha\alpha$Base through BRAPI.

% % Th to encourage public (and perhaps private?) breeding programs to contribute and use data to . We want to implement a genomic prediction tool for alfalfa breeders that will also design idealized mating schemes .

% I intend to continue to foster collaborations built by Dr. Don Viands and his forage breeding research team, while hoping to expand them with new research groups, such as that of Animal Nutrition (Cornell), Ian Ray (NMSU), Charlie Brummer (UC Davis), Maria Monteros (Noble Foundation), and Debby Samac (USDA ARS at UMN, and the new forage breeder they are currently looking to hire). The forage breeding program at Cornell consists of a strong team, whose knowledge and experience will be invaluable to implementing a successful GS breeding program. Additional personnel will be brought on as necessary, including a technician to operate and manage all HTP equipment and activities.



% If proven useful, I intend to establish collaborations to implement the strategies outlined here in other forage programs in the US and abroad.    

% Finally, I recognize that the forage breeding project currently focuses on many forages other than alfalfa. It is my intent to keep those projects, while incorporating new technologies for genetic gain in these crop. My focus on alfalfa is driven by the need for reliable GS methods in a polyploid crop that must be bred at the population level. I am open to new species as well, especially if a graduate student has a desire to investigate such species.  



% The third grant proposal will be to unify breeding efforts. Genotype by genotype interactions. 

% Field establishment 


% Other long term goals include:

% Collaborate with the Agro-ecology biochemistry.  CRISPR will be used to change both the Rhizobium as well as the plant. 

% Collaboration with the animal nutrition department. sciences


% A technician will be recruited to operate the HTP platform. 

% Implementation of marker-based selection has also lagged because large numbers of individuals must be genotyped to avoid inbreeding depression in future generations. 






\end{document}
