\documentclass[11pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[backend=bibtex, style=authoryear]{biblatex}

% \addbibresource{~/Dropbox/TAGreview/PhDreferences.bib}

\textwidth=470pt
\oddsidemargin=0pt
\topmargin=0pt
\headheight=0pt
\textheight=680pt
\headsep=0pt

\title{Statement of Research}


\author{Nicholas Santantonio}
\date{\today}


\newcommand{\gxe}{G$\times$E}
\newcommand{\gxg}{G$\times$G}


\begin{document}

\section*{\centering Statement of Teaching}
\begin{center} Nicholas Santantonio \end{center}
% \subsection*{Introduction}

\noindent Plant breeders have traditionally been generalists, combining genetics with a range of plant sciences to identify farmers' needs and produce products to meet those needs. However, the range of skills required in the field is rapidly increasing. Competitive plant breeders are now expected to be proficient in statistics, programming, bioinformatics and computational biology along with the traditional skills in physiology, pathology, agronomy and genetics. As educators, we must decide how to update our curriculum to best prepare students for careers in the post-genomics era of digital agriculture. Should we encourage a mile wide and an inch deep approach? Or should we recognize that the discipline has grown so wide in breadth that we must instead encourage specialization? 

Both will be needed for integrating genomics and digital agriculture into modern breeding programs. Generalists may require instruction in team management, while specialists may need more rigorous courses in areas previously considered outside plant breeding, such as engineering and computer science. Specialists will be vital for collecting and storing large amounts of data that will be used to model and predict crop growth and development. The generalists will coordinate those activities to lead a team that produces new varieties to meet farmers' needs. %The curriculum has lagged in preparing individuals for these positions. 

% building sensing systems to collect large amounts of data, storing that data in a database, and accessing that data to  This information will then be used to make informed decisions about breeding goals and design optimal mating schemes. 

% As educators, we must actively update our curriculum to prepare students to be competitive in the workplace if Cornell is to continue to be a leader in the plant breeding community. This is especially true today, where plant breeding has seen a rapid and unprecedented paradigm shift in recent years away from tinkering with single genes and toward whole genome engineering. 

Due to the wide breadth of the plant breeding discipline, it may be pertinent to develop multiple paths of instruction. While I believe all future scientists will require some degree of graduate level statistics and computational coursework, a more structured series of quantitative and computational courses would benefit students seeking specialization. Students who chose this path would come out of graduate school with the comprehension and ability to effectively use the latest computational techniques in plant breeding, including the use of genome-wide information for decision making.




% At the undergraduate level, plant breeding courses at NMSU are rather limited. Heightened interest in food systems has drawn many young people to seek out information about how crops have been genetically manipulated by humans. An earlier course at the 200 level would be ideal to pique interest and motivate students to pursue plant breeding as a profession. However, there is relatively little infrastructure to foster such an interest once it is seeded. A stronger undergraduate curriculum in new and traditional plant breeding technologies and quantitative genetics would increase enrollment in the Plant Science and Genetics degree paths and add to the department's value to the university.

% While single- and oligo-gene mendelian triats are abundant, why should students be limited to this type of genetic variation?  

\subsection*{Teaching Philosophy}

It is important for students to be exposed to plant breeding ideas from multiple perspectives, and they must given the opportunity to demonstrate critical thinking on different levels. While some students may show analytical thinking and synthesis on exams, others may shine in more hands-on projects. Most mathematical and computational learning occurs through doing, not through watching. The lecture is important to present material in a concise structured manner, but concepts are cemented when the student can reconstruct the ideas on their own time.

% As a student, my most memorable courses were those that had a semester-long project that was presented to the class at the end of the course. 

Longer term projects provide students the opportunity to practice and apply concepts in a more autonomous environment, and are invaluable for assessing comprehension and critical thinking. Regularly assigned homework and hands-on labs are also crucial for estimation of the pace and overall understanding of the material presented so that adjustments can be made where necessary. 

The greenhouse and field facilities on campus provide good opportunities to get students out of the classroom to see plant breeding in practice. Public databases provide resources where students can find datasets to work with and computer simulations are useful tools to evaluate comprehension, where in order to simulate a system correctly, the student must understand that system well.

% Hands-on labs are also useful for evaluating individual students, as it becomes easier to spot students that may need extra attention if they are struggling in labs as well as on homework assignments. 

% The greenhouse and field aspects of plant breeding are difficult to present in lecture courses. One of the benefits of having greenhouse and field space on campus is that students can be brought out of the classroom to see what breeding looks like in practice. The computational aspects of plant breeding can be incorporated into the curriculum through homework assignments, labs and semester long projects. 

% Public databases provide invaluable resources where students can find data sets for use in projects that teach them how to work with real, and often large data sets. 


All courses I instruct would contain a term project of relevant complexity to augment exams, in-class labs and homework assignments. For quantitatively oriented courses, these projects would consist of a hands-on computational component, where students either find a dataset or use their own data to explore the ideas covered in the course. They would then be asked to present their results in written and oral formats that mirror typical scientific communication. Projects may be team oriented to promote collaborative skills and project management. For courses without a significant quantitative aspect, term projects may be formulated as a research proposal.%, where the students would be expected to construct a hypothesis, give background information on merit and propose a way to ask their question while also considering potential pitfalls. %These proposals would be guided throughout the semester to ensure students can 

% Any course related to quantitative genetics should have a computational aspect to it, 

I also intend to incorporate more perspectives on how domestication, selection and seed systems work in different cultures through guest lectures and case studies. Review of how seed systems function in other cultures will cultivate discussion on the benefits and consequences of these systems relative to the US and western European systems. 

\subsection*{Courses}

I propose to teach one 400 level undergraduate/graduate course a year, as well as one advanced graduate course every (other) year to meet the 25\% teaching requirement of the position. The Plant Breeding (AGRO462) course may make a good place to introduce more quantitative ideas to prospective plant breeding students. Additionally, I would like to teach a more advanced quantitative genetics course to complement other quantitative courses, such as AGRO610 and AGRO670.
% such as Biometrical Genetics and Plant Breeding (AGRO670) and Advanced Crop Breeding (AGRO610).

% If I were to instruct AGRO462, the course would shift from 20th century breeding concepts in the beginning toward aspects of 21st Century plant breeding and digital agriculture in the latter half of the course. The course would be taught from a historical perspective, allowing the natural progression of plant breeding over the past century to be a teaching tool for how we arrived at the methods we use today. By the end of the semester, students will be able to think critically about how breeding technology changes through time have impacted the discipline and the societal view thereof. 

If I were to instruct AGRO462, the course would include more focus on $21^\text{st}$ century plant breeding. The class would be taught from a historical perspective, allowing the natural progression of plant breeding over the past century to be a teaching tool for how we arrived at the methods we use today. Instead of a textbook, primary literature would be used to introduce key concepts. By the end of the semester, students will be able to think critically about how breeding technology changes through time have impacted the discipline and the societal view thereof. 

I would incorporate more hands-on computational labs and a term project to supplement the literature review framework. Computational labs would be used to augment student understanding of course material, in which students would use available computational tools to analyze example datasets with a focus on interpretation of results. Students would be able to demonstrate critical thinking of plant breeding methods and ideas, with the ability to synthesize when given new plant systems or breeding goals. 

% This would include emphasis on the digital agriculture aspect of plant breeding toward the second half of the semester. 

As for the graduate level course, I would like to offer a quantitative genetics theory or statistical genomics course that would include a weekly computational lab and a student term project which would be developed throughout the semester. Starting with basic probability theory and the single locus model, the course would advance through genome-wide association, genomic prediction and selection, as well as selection theory and breeding program optimization. 

Concepts in class would be reinforced through labs and assignments that require students to write their own software to solve plant breeding problems. Assignments would be required to be submitted as typed documents in Markdown, \LaTeX\ or similar format, to expose students to more effective modes of mathematical communication outside of Microsoft Office. The term project might then consist of groups of 2-4 students finding a genotype-phenotype dataset, and working to develop a genotype to phenotype map and assess genomic predictability throughout the semester.

%Proposal outlines would be submitted half way through the semester to evaluate project appropriateness, and students would be encouraged to seek input from the instructor throughout the semester. Late semester labs would be used for working on projects, where code readability and reproducibility would be emphasized.% with a focus on interpretation of results.% instead of method theory.

% The course topic is flexible to the needs of the section, however, it is my observation that an introductory plant breeding quantitative/computational course would be useful to help prepare students for further quantitative study. This course might also provide some necessary skills to those who choose not to follow a quantitative path, without the rigorous mathematical detail of the PLBRG7160 quantitative genetics course.


\subsection*{Curriculum}

Currently, most life science students do not acquire in-depth statistics and programming skills until graduate school, impeding their progress while they learn to grapple with these new languages. Moving forward, I would like to work with faculty in PES, statistics, engineering and computer science to build a quantitative/computational genetics undergraduate curriculum. Genetics is a vast field, and it is imperative that students get exposure to and training in the rapidly changing environment in which they will soon be seeking jobs. While ``single-gene'' genetics is still an important field, more and more the community is focused on the complex ``omics'' network that is the foundation of complex organisms. Students in essentially all sub-fields of genetics will need to be able to deal with large datasets, using tailored algorithms to make inferences, predict the unobserved, and guide decision making. Dealing with large data requires skills in programming, linear algebra, statistics and machine learning. 


% I would like to work with faculty within PES and outside the department in statistics and computational biology to build a. Lower level courses would be aimed at recruiting students' interests, while more advanced courses would be targeted to those who choose to specialize.


% This would involve reducing undergraduate emphasis on chemistry, biochemistry and molecular genetics, while increasing math, computer science, and statistics.




% the ability to program in a flexible scripting language, understand complex mathematical and statistical concepts, as well as familiarity with common algorithms used for such purposes. 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


% Algorithms for quantitative genetics, which would cover the basic algorithms used to solve for unknown parameters and solutions. Students would learn how to build expectation maximization, Markov Chain Monte Carlo (e.g. Metropolis-Hastings) swarm intelligence (e.g. ant colony optimization), and genetic algorithms common place in much of the software they will use throughout their careers. While this will be insufficient for these students to write their own computationally efficient software, it will provide a basis for students to understand how more optimized software works from a theoretical perspective.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


% There are several courses offered by the plant breeding section that may need instructors in the near future, including PLBRG2010 and PLBRG4030. I worked closely with Professor Susan McCouch as a TA for PLBRG2010 in fall of 2017, and would be honored to continue as that course's instructor. My background would also be appropriate to teach the plant breeding staples of PLBRG4030 while incorporating a more hands-on computational focus to further prepare students for study of plant breeding on a quantitative level. 


% The term project will be an exploration into a ``scientific, societal, or technological challenge'' related to plant breeding in the form a research proposal. Students will submit outlines and meet with the instructor mid-semester to evaluate the appropriateness of the proposed topic. ``Crop of the Day'' presentations will be assigned during the first week, where each student will choose a crop on which to present information about the origin, biology and breeding methods/goals of the crop, to be given at the beginning of each class.



% The term project would be an investigation into a quantitative problem of their choosing, where the students will be free to use existing data, synthesize their own data and/or simulate a genetic system using tools they will be learning and writing along the way. 


% Finally, I want to work with Assistant Professor Kelly Robbins and faculty from biometry and computational biology to produce another entry level graduate or cross-listed course that will aim to prepare students for quantitative and computational genetics. Currently, there is a single graduate level course on quantitative genetics offered by the Plant Breeding and Genetics Section, under the instruction of Dr. Kelly Robbins. That course would benefit greatly from a introductory course, such that more time can be dedicated to an in-depth mathematical treatment of quantitative genetics theory for students who would like to specialize in quantitative genetics. This introductory course would provide the background theory, coding and application of known methods, while reserving focus on mathematical proofs and Bayesian methods for the advanced course. I envision essentially all plant breeding graduate students to take this course to gain the basic skills, while the advanced class would be available for those who wish to pursue the discipline further. Covered topics may include review of probability, single gene model, genetic variances and covariances, linear regression, linear mixed models, genome wide association and genomic prediction.
 % who do not wish to pursue the discipline further, but need the basic tools. The introductory QG course will focus on coding and application of known methods

% The course
% At the semester conclusion, students will demonstrate mastery of regression and linear mixed models to solve GWAS and GS problems. Students will be able to build design matrices and fit linear mixed models to solve for random effects with known genetic covariances in a computational environment. Students will have developed the computational tools to solve many of the problems they may encounter in their graduate research, while also developing a foundation from which to build new ideas to tackle new problems. 
% My expertise is quantitative in nature, and I would want to incorporate more of the quantitative element to the course so that students are exposed to what the plant breding field looks like today, and what it may look like in the future.

% I propose to teach two courses, one of either PLBRG2010 or PLBRG4030, as well as a new cross-listed, ``Introduction to Quantitative Genetics in Plant Breeding'' 4000 level course. The new course would prepare senior undergraduates and first year graduate students for further study of quantitative genetics and would cover some background material currently needed for PLBRG7160. I would like to work with Assistant Professor Kelly Robbins, as well as faculty in Biometry and Computational biology to determine what material should be covered in this cross-listed course, therefore allowing PLBRG7160 to be an advanced course that provides more rigorous mathematical detail to quantitative genetics.




% The new course would prepare senior undergraduates and first year graduate students for further study of quantitative genetics and would cover some background material currently needed for PLBRG7160. I would like to work with Assistant Professor Kelly Robbins, as well as faculty in Biometry and Computational biology to determine what material should be covered in this cross-listed course, therefore allowing PLBRG7160 to be an advanced course that provides more rigorous mathematical detail to quantitative genetics.

% % \section*{(PLBRG2010)}
% % \section*{PLBRG 2XXX - Genome to Table: Technological advances in plant breeding}
% \subsection*{PLBRG 2010 - Plants, Genes, and Global Food Production}



% % Topics covered in the course would include a brief review of Mendelian genetics, domestication, genetic diversity, mating systems, the Green Revolution, biotechnology, intellectual property, gene mapping, genetics of complex traits, and introduction to new plant breeding technologies such as genomic selection and high throughput phenotyping.

% %The course will introduce students to plant breeding as a discipline and cover the basics of domestication, genetic diversity, and mating systems before shifting to a quantitative perspective to introduce students to these topics such that they may seek further quantitative and computational instruction before they graduate. %Topics covered in the course would include a brief review of Mendelian genetics, domestication, genetic diversity, mating systems, the Green Revolution, biotechnology, intellectual property, gene mapping, genetics of complex traits, and introduction to new plant breeding technologies such as Genomic selection and high throughput phenotyping.

% % \textbf{Objective:}  Students will be able to express how breeding technology changes through time have impacted the discipline and the societal view thereof. 

% % Student performance will be evaluated by three exams, regularly assigned homework, a term project, and a single a ``Crop of the Day'' presentation. The term project will be an exploration into a scientific, societal, or technological challenge related to plant breeding a in the form a research proposal. Students will submit outlines and meet with the instructor mid-semester to evaluate the appropriateness of the proposed topic. ``Crop of the Day'' presentations will be assigned during the first week, where each student will choose a crop on which to present information about the origin, biology and breeding methods/goals of the crop, to be given at the beginning of each class.


% \subsection*{PLBRG 4030 - Genetic Improvement of Crop Plants}


% The course will introduce students to plant breeding as a discipline and cover the basics of domestication, genetic diversity, and mating systems before shifting to a quantitative perspective to introduce students to these topics such that they may seek further quantitative and computational instruction before they graduate. %

% \textbf{Objective:} Students will be able demonstrate critical thinking of plant breeding ideas, with the ability to synthesize when given new information or situations. Students will be able to express how breeding technology changes through time have impacted the discipline and the societal view thereof. 

% Student performance will be evaluated by three exams, regularly assigned homework, a term project, and a single a ``Crop of the Day'' presentation. The term project will be an exploration into a scientific, societal, or technological challenge related to plant breeding a in the form a research proposal. Students will submit outlines and meet with the instructor mid-semester to evaluate the appropriateness of the proposed topic. ``Crop of the Day'' presentations will be assigned during the first week, where each student will choose a crop on which to present information about the origin, biology and breeding methods/goals of the crop, to be given at the beginning of each class.


% The course will introduce students to the basics of plant breeding from a historical perspective and transition to the quantitative and computational aspects of the rapidly evolving field. 

% \begin{itemize}

% 	\item A brief review of Mendelian genetics (Mendel, Morgan, etc.)
% 	\begin{itemize}
% 		\item 1 and 2 gene inheritance
% 		\item chromosomes, linkage
% 	\end{itemize} 

% 	\item Domestication and genetic diversity (Nikolai Vavilov, Cary Fowler, etc.)
% 	\begin{itemize}
% 		\item Centers of origin 
% 		\item Domestication traits
% 		\item Gene banks: seeds vs clonal species
% 	\end{itemize} 

% 	\item Mating systems and ploidy (with pollination lab)
% 	\begin{itemize}
% 		\item Inbred crops
% 		\item Outcrossing crops
% 		\item Hybrids  
% 	\end{itemize} 

% 	\item Green Revolution (Norman Borlaug, etc.)
% 	\begin{itemize}
% 		\item Impact of inorganic nitrogen
% 		\item Asia
% 		\item Africa
% 	\end{itemize} 

% 	\item Biotechnology \& Intellectual Property
% 	\begin{itemize}
% 		\item Traditional GMOs (with a contrast to mutagenesis)
% 		\item Gene editing - GMO or not?
% 		\item PVP and Patents 
% 	\end{itemize}

% 	\item Locating genes (with linkage mapping lab)
% 	\begin{itemize}
% 		\item Linkage mapping 
% 		\item GWAS (intro)
% 	\end{itemize} 

% 	\item Complex traits (Fisher, etc.)
% 	\begin{itemize}
% 		\item Additivity
% 		\item Dominance
% 	\end{itemize} 

% 	\item Introduction to new(er) technologies
% 	\begin{itemize}
% 		\item Marker-assisted selection (MAS)
% 		\item Genomic selection (GS)
% 		\item High throughput phenotyping (HTP)
% 	\end{itemize} 

% \end{itemize}




% \subsection*{PLBRG 4$\cdots$ - Introduction to Quantitative Genetics for Plant Breeding }

% Finally, I want to work with Assistant Professor Kelly Robbins and faculty from biometry and computational biology to produce another entry level graduate or cross-listed course that will aim to prepare students for quantitative and computational genetics. Currently, there is a single graduate level course on quantitative genetics offered by the Plant Breeding and Genetics Section, under the instruction of Dr. Kelly Robbins. That course would benefit greatly from a introductory course, such that more time can be dedicated to an in-depth mathematical treatment of quantitative genetics theory for students who would like to specialize in quantitative genetics. This introductory course would provide the background theory, coding and application of known methods, while reserving focus on mathematical proofs and Bayesian methods for the advanced course. I envision essentially all plant breeding graduate students to take this course to gain the basic skills, while the advanced class would be available for those who wish to pursue the discipline further. Covered topics may include review of probability, single gene model, genetic variances and covariances, linear regression, linear mixed models, genome wide association and genomic prediction.
%  % who do not wish to pursue the discipline further, but need the basic tools. The introductory QG course will focus on coding and application of known methods

% The course would include a weekly computational lab and a paired student term project which will be developed throughout the semester. All homework assignments will contain both theoretical and computational problems, that must be submitted as typed documents in Markdown, \LaTeX\ or similar format. The term project will be an investigation into a quantitative genetics problem of their choosing, where the students will be free to use existing data, synthesize their own data and/or simulate a genetic system using tools they will be learning and writing along the way. Proposal outlines will be submitted half way through the semester to evaluate project appropriateness, and students will be encouraged seek input from the instructor throughout the semester. Late semester weekly labs will be used for working on projects, where code readability and reproducibility will be emphasized.

% At the semester conclusion, students will demonstrate mastery of regression and linear mixed models to solve GWAS and GS problems. Students will be able to build design matrices and fit linear mixed models to solve for random effects with known genetic covariances in a computational environment. Students will have developed the computational tools to solve many of the problems they may encounter in their graduate research, while also developing a foundation from which to build new ideas to tackle new problems. 


 % The term project will consist of students finding a genotype-phenotype dataset, and working to develop a genotype to phenotype map and assess genomic predictability throughout the semester, .

% \textbf{Objective:} 
% \begin{itemize}

% 	\item Review of probability
% 	\begin{itemize}
% 		\item Random variables
% 		\item Probability density functions (PDF) and probability mass functions (PMF)
% 		\item Central limit theorem and the normal distribution
% 	\end{itemize} 

% 	\item Single gene model 
% 	\begin{itemize}
% 		\item Additive
% 		\item Dominance 
% 		\item Effect of inbreeding
% 	\end{itemize} 

% 	\item Genetic variances and covariance
% 	\begin{itemize}
% 		\item Additive and dominance covariance
% 		\item Calculating expectations from a pedigree 
% 		\item Calculating realized covariances from marker scores
% 		\item Necessity of marker imputation
% 	\end{itemize} 

% 	\item Linear regression
% 	\begin{itemize}
% 		\item Parameters and likelihood
% 		\item Numerical and categorical predictors 
% 		\item Ordinary least squares
% 		\item Connection to ANOVA/ANCOVA
% 	\end{itemize} 

% 	\item The linear mixed model
% 	\begin{itemize}
% 		\item Random versus fixed effects
% 		\item Known and estimated covariances
% 		\item Expectation Maximization (EM) algorithm for solving mixed model equations 
% 	\end{itemize} 

% 	\item Genome wide association
% 	\begin{itemize}
% 		\item Linkage Disequilibrium
% 		\item Population structure
% 		\item QQ plots
% 	\end{itemize} 

% 	\item Genomic Prediction
% 	\begin{itemize}
% 		\item Evaluation of model fit 
% 		\item Genomic prediction of unobserved individuals
% 		\item k-fold cross-validation 
% 	\end{itemize} 

% \end{itemize}



% \subsection*{Undergraduate degree in computational plant science}

% In the future I seek to help refine the undergraduate degree program in plant science oriented to quantitative genetics and computational biology. The undergraduate degree would have a heavy focus on linear algebra, programming, statistics, machine learning and operational research. 


% Bachelors in unified breeding

% \begin{itemize}

% 	\item Core requirements
% 	\begin{itemize}	
% 		\item Math (linear algebra, matrix calculus)
% 		\item Statistics (prob and Stat theory)
% 		\item Programming (Python / C++ / etc.)
% 		\item Machine learning
% 		\item Operational research (OR dept?)
% 		\item Biology
% 		\item Chemistry
% 	\end{itemize} 

% 	\item PLBR requirements
% 	\begin{itemize}
% 		\item Plant Genetics (PLBRG 2250, Mazourek)
% 		\item Genetic Improvement of Crop Plants (PLBRG 4030, Gore/Reisch)
% 		\item Introduction to quantitative genetics (PLBRG 40XX , Santantonio)
% 		% \item New age of Plant and Animal breeding (GS etc...)
% 	\end{itemize} 

% 	\item Plant Science Electives (must pick two)
% 	\begin{itemize}
% 		\item Pathology 
% 		\item Physiology
% 	\end{itemize} 

% 	\item Management
% 	\begin{itemize}
% 		\item Course about building a successful team
% 	\end{itemize} 

% \end{itemize}
% \subsection*{Conclusion}

\medskip

Genomics and digital agriculture is only just starting to change the landscape of food production. Quantitative skills are one of the specializations imperative in plant science, and NMSU must be at the forefront of preparing the best individuals to usher in this new era.

% along with engineering, computer science, operational research and the traditional disciplines
\end{document}
